---
title: "Getting started with sharpapi"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with sharpapi}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
# Every chunk that would call the API is eval = FALSE: building this vignette
# must not require a key or a network connection.
```

## What this package does

`sharpapi` wraps the [SharpAPI](https://sharpapi.io) REST API, which normalizes
real-time sports betting odds from 45+ sportsbooks into one schema. Four
functions cover the common cases:

| Function | Endpoint | Minimum tier |
|---|---|---|
| `sharpapi_sports()` | `/sports` | any |
| `sharpapi_odds()` | `/odds` | any |
| `sharpapi_arbitrage()` | `/opportunities/arbitrage` | Hobby |
| `sharpapi_ev()` | `/opportunities/ev` | Pro |

Each returns a plain data frame, so the result goes straight into whatever you
normally use — `dplyr`, `data.table`, base subsetting.

## Setting your key

Every call needs an API key. A free tier is available at
<https://sharpapi.io>. The package reads credentials from the `SHARPAPI_KEY`
environment variable rather than function arguments. Avoid storing real keys
in source files:

```{r, eval = FALSE}
Sys.setenv(SHARPAPI_KEY = "your-key-here")
```

For anything beyond a one-off session, put it in `~/.Renviron` instead
(`usethis::edit_r_environ()` opens the file) so it is set for every session and
stays out of your project directory:

```
SHARPAPI_KEY=your-key-here
```

If the variable is unset, every function stops with a message telling you so
rather than making an unauthenticated request.

## Which sports are live right now

```{r, eval = FALSE}
library(sharpapi)

sports <- sharpapi_sports()
head(sports[order(-sports$live_count), c("id", "name", "event_count", "live_count")])
```

```
#>          id       name event_count live_count
#> 1    soccer     Soccer         812         37
#> 2  baseball   Baseball          51          4
```

## Fetching odds

`sharpapi_odds()` passes any named argument through to the API as a filter.
Useful ones are `sport`, `league`, `sportsbook`, `market_type`, `is_live` and
`limit`:

```{r, eval = FALSE}
odds <- sharpapi_odds(sport = "baseball", market_type = "moneyline", limit = 500)

str(odds[1, c("sportsbook", "home_team", "away_team", "selection", "odds_american")])
```

```
#> 'data.frame': 1 obs. of  5 variables:
#>  $ sportsbook   : chr "pinnacle"
#>  $ home_team    : chr "Yankees"
#>  $ away_team    : chr "Red Sox"
#>  $ selection    : chr "Yankees"
#>  $ odds_american: int -145
```

Because the result is a data frame, comparing one market across books is
ordinary R:

```{r, eval = FALSE}
best <- odds[odds$event_id == odds$event_id[1] & odds$selection == "Yankees", ]
best[order(-best$odds_decimal), c("sportsbook", "odds_american", "odds_decimal")]
```

## Finding opportunities

`sharpapi_arbitrage()` (Hobby tier and above) returns cross-book arbitrage. The
individual sides live in a nested `legs` column rather than as top-level
fields:

```{r, eval = FALSE}
arb <- sharpapi_arbitrage(sport = "baseball")
arb[, c("event_id", "market_type", "profit_percent")]

# the two sides of the first opportunity
arb$legs[[1]]
```

`sharpapi_ev()` (Pro tier and above) returns positive expected value bets,
priced against the sharp reference books:

```{r, eval = FALSE}
ev <- sharpapi_ev(sport = "baseball")
head(ev[order(-ev$ev_percentage), c("sportsbook", "selection", "ev_percentage")])
```

A key below the required tier raises an HTTP error rather than returning an
empty frame, so wrap these in `try()` if you are probing what a key can reach.

## Trying it without a key

If you want to work through the data shapes before signing up, sample
snapshots (MLB and 2026 World Cup) are published as CSVs at
<https://github.com/Sharp-API/SharpAPI-Sample-Data>. The snapshots use SharpAPI odds
field names and provide offline examples for exploring the data. Live responses
may include additional fields or nested objects:

```{r, eval = FALSE}
url <- paste0(
  "https://raw.githubusercontent.com/Sharp-API/",
  "SharpAPI-Sample-Data/main/data/mlb_odds_snapshot.csv"
)
odds <- read.csv(url)
```

## Errors

HTTP errors are raised by `httr2` with the response status and API message.
Common status codes include:

| Status | Meaning |
|---|---|
| 401 | key missing or invalid |
| 403 | key is valid but below the tier this endpoint requires |
| 429 | rate limit for your tier exceeded |
